Why is it that statistical tests for residuals are not widely used? An explanation using visual inference.

### 2023-02-15 paper_comparison/paper_comparison/paper_comparison.Rmd
clarify the power and the simulated conventional power

\noindent where $M$ is a given model that generates the data, $diag(.)$ is the diagonal matrix constructed from the diagonal elements of a matrix,  $\boldsymbol{X}_a = (\boldsymbol{1},  \boldsymbol{X})$ is the matrix of regressors including the intercept used in the regression equation, $\sigma^2\boldsymbol{I}$ is the assumed variance of the error term when $H_0$ is true, $\boldsymbol{V}$ is the actual variance of the error term, and $\boldsymbol{\mu}_z = \boldsymbol{R}_a\boldsymbol{Z}\boldsymbol{\beta}_z$ is the expected values of the residuals with $\boldsymbol{Z}$ be any variables leave out by the regression equation and $\boldsymbol{\beta}_z$ be the corresponding coefficients. An expectation is taken because $E_{sample}$ depends on the regressors $X$ and we want the effect size $E$ to be the same for the same data generating process. It is difficult to solve the effect size $E$ analytically, so we estimate it with the average of $E_{sample}$ over a very large sample size.

@@

In the following analysis, lineups with uniform fitted values will be the focus. Visual patterns are more likely to be revealed under a uniform distribution. Additionally, we have collected extra evaluations on these lineups, which will result in more reliable analysis. Analysis of lineups with other fitted value distributions can be found in Section \ref{effect-of-fitted-value-distributions}.

@@

To compare the power of visual test and conventional test, we pick RESET test (`resettest`) and Breusch-Pagan test (`bptest`) from the R package `lmtest`, and Shapiro-Wilk test (`shapiro.test`) from the built-in R package `stats`. Among them, RESET test is the only exact and appropriate test in this scenario. Both the Breusch-Pagan test and the Shapiro-Wilk test are approximate and inappropriate tests. Their estimated power is shown in Figure \ref{fig:polypower}. To set up the RESET test, we include different powers of fitted values as proxies. According to @ramsey_tests_1969, there are no general rules for the power of the fitted values needed by the RESET test, but it finds power up to four is usually sufficient. Thus, we follow this guideline to conduct the RESET test. For the Breusch-Pagan test, the choice of regressors in the auxiliary regression is left to the user [@breusch_simple_1979]. But as @waldman1983note suggested, it is a good choice for the set of auxiliary regressors in the Breusch-Pagan test be the same as the White test. Thus, we include both $\boldsymbol{x}$ and $\boldsymbol{x}^2$ in the auxiliary regression.

>>>>

\noindent where $M$ is a given model that generates the data, $diag(.)$ is the diagonal matrix constructed from the diagonal elements of a matrix,  $\boldsymbol{X}_a = (\boldsymbol{1},  \boldsymbol{X})$ is the matrix of regressors including the intercept used in the regression equation, $\sigma^2\boldsymbol{I}$ is the assumed variance of the error term when $H_0$ is true, $\boldsymbol{V}$ is the actual variance of the error term, and $\boldsymbol{\mu}_z = \boldsymbol{R}_a\boldsymbol{Z}\boldsymbol{\beta}_z$ is the expected values of the residuals with $\boldsymbol{Z}$ be any variables leave out by the regression equation and $\boldsymbol{\beta}_z$ be the corresponding coefficients. A conditional expectation is taken because $E_{sample}$ depends on the regressors $\boldsymbol{X}$ and we want the effect size $E$ to be the same for the same data generating process. It is difficult to solve the effect size $E$ analytically, so we estimate it with the average of $E_{sample}$ over a very large sample size.

@@

In the following analysis, lineups with uniform fitted values will be the focus. Visual patterns are more likely to be revealed under a uniform distribution. Additionally, we have collected extra evaluations on these lineups, which will result in a more reliable analysis. Analysis of lineups with other fitted value distributions can be found in Section \ref{effect-of-fitted-value-distributions}.

@@

To compare the power of visual tests and conventional tests, we simulate 200000 samples evenly from the models used in the human subject experiments, and record the conventional $p$-values produced from every sample. We pick RESET test (`resettest`) and Breusch-Pagan test (`bptest`) from the R package `lmtest`, and Shapiro-Wilk test (`shapiro.test`) from the built-in R package `stats`. Among them, RESET test is the only exact and appropriate test in this scenario. Both the Breusch-Pagan test and the Shapiro-Wilk test are approximate and inappropriate tests. Their estimated power is shown in Figure \ref{fig:polypower}. To set up the RESET test, we include different powers of fitted values as proxies. According to @ramsey_tests_1969, there are no general rules for the power of the fitted values needed by the RESET test, but it finds power up to four is usually sufficient. Thus, we follow this guideline to conduct the RESET test. For the Breusch-Pagan test, the choice of regressors in the auxiliary regression is left to the user [@breusch_simple_1979]. But as @waldman1983note suggested, it is a good choice for the set of auxiliary regressors in the Breusch-Pagan test be the same as the White test. Thus, we include both $\boldsymbol{x}$ and $\boldsymbol{x}^2$ in the auxiliary regression.

To get an accurate estimate of the power of conventional tests, we simulate 200000 samples evenly from the models used in the human subject experiments, and record the $p$-values produced by every sample.

### 2023-02-15 paper_comparison/paper_comparison/paper_comparison.tex
clarify the power and the simulated conventional power

\noindent where \(M\) is a given model that generates the data,
\(diag(.)\) is the diagonal matrix constructed from the diagonal
elements of a matrix,
\(\boldsymbol{X}_a = (\boldsymbol{1}, \boldsymbol{X})\) is the matrix of
regressors including the intercept used in the regression equation,
\(\sigma^2\boldsymbol{I}\) is the assumed variance of the error term
when \(H_0\) is true, \(\boldsymbol{V}\) is the actual variance of the
error term, and
\(\boldsymbol{\mu}_z = \boldsymbol{R}_a\boldsymbol{Z}\boldsymbol{\beta}_z\)
is the expected values of the residuals with \(\boldsymbol{Z}\) be any
variables leave out by the regression equation and
\(\boldsymbol{\beta}_z\) be the corresponding coefficients. An
expectation is taken because \(E_{sample}\) depends on the regressors
\(X\) and we want the effect size \(E\) to be the same for the same data
generating process. It is difficult to solve the effect size \(E\)
analytically, so in the analysis, we estimate it with the average of
\(E_{sample}\) over a very large sample size.

>>>>

\noindent where \(M\) is a given model that generates the data,
\(diag(.)\) is the diagonal matrix constructed from the diagonal
elements of a matrix,
\(\boldsymbol{X}_a = (\boldsymbol{1}, \boldsymbol{X})\) is the matrix of
regressors including the intercept used in the regression equation,
\(\sigma^2\boldsymbol{I}\) is the assumed variance of the error term
when \(H_0\) is true, \(\boldsymbol{V}\) is the actual variance of the
error term, and
\(\boldsymbol{\mu}_z = \boldsymbol{R}_a\boldsymbol{Z}\boldsymbol{\beta}_z\)
is the expected values of the residuals with \(\boldsymbol{Z}\) be any
variables leave out by the regression equation and
\(\boldsymbol{\beta}_z\) be the corresponding coefficients. A
conditional expectation is taken because \(E_{sample}\) depends on the
regressors \(\boldsymbol{X}\) and we want the effect size \(E\) to be
the same for the same data generating process. It is difficult to solve
the effect size \(E\) analytically, so we estimate it with the average
of \(E_{sample}\) over a very large sample size.
